Life in the canopy: Phylogenetic filtering and morphological canalization in epiphytic ferns

### 2018-03-03 tables.Rnw
Add tables.Rnw to output only tables

\documentclass{article}
\date{}

@@

% Set line spacing to 1.5
\usepackage{setspace}
\onehalfspace

@@

% Set unit and abbreviations formatting
\input{newphyt_style}

@@

<<setup, echo=FALSE>>=
# setup knitr
library(knitr)
opts_chunk$set(echo=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(warning=FALSE)
@

<<setup_tables>>=
library(xtable)
source("bin/latex_functions.R")
# load results objects produced by running manuscript.Rnw
load("results.RData")
@

\noindent \textbf{\textit{New Phytologist} Supporting Information} \\
\\
\noindent Article title: Life in the canopy: phylogenetic filtering and morphological canalization in epiphytic ferns \\
\noindent Authors: Joel H. Nitta, James E. Watkins, Jr., Ravahere Taputuarai and Charles C. Davis \\
\noindent Article acceptance date: someday soon \\
\\
\noindent The following Supporting Information is available for this article: \\

\noindent \textbf{Methods S1} We performed alternative tests for correlation of growth habit with functional traits using phylogenetic generalized linear mixed models (PGLMMs). Models were constructed with each trait as the response variable in turn dependent on growth habit, with phylogeny as a random effect. For binary (gametophyte) traits we used the \function{binaryGLMM} function in the \package{ape} package \citep{Paradis2004}. For quantitative (sporophyte) traits, we used the \function{MCMCglmm} function in the R package of the same name \citep{Hadfield2010}. Priors were set to an inverse-Gamma distribution with shape and scale parameters equal to \num{0.01} \citep{Hadfield2010a}. The MCMC algorithm was run for \num{5000000} iterations with a burnin of \num{1000} and thinning every \num{500} iterations.

@@

% Add S to table number and reset counter
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

% Table S1 (supp_table:PGLMM_sporo)
<<PGLMM, results="asis">>=
# read in results from running PGLMM on cluster, clean up dataframe
pglmm.results <- read.csv("data/mcmcglmm_habit_all_results.csv")
pglmm.results <- pglmm.results[pglmm.results$Effect == "habit", ]
pglmm.results$Effect <- NULL
pglmm.results$Model <- gsub(" ~ habit", "", pglmm.results$Model)
pglmm.results$Model <- tolower(pglmm.results$Model)
rownames(pglmm.results) <- pglmm.results$Model
pglmm.results$Model <- NULL

@@

% Table S2
\clearpage
<<binaryPGLMM, results="asis">>=
bpglmm.results <- round_df(bpglmm.results)
bpglmm.results <- embolden_p(bpglmm.results, col.select=c("sigmap","pvalue"))

@@

% Table S3
\clearpage
<<climate_ancova, results="asis">>=
ancova.results <- round_df(ancova.results)
ancova.results <- embolden_p(ancova.results, col.select=c("pval.intercept","pval.slope"))
part_label <- add_part_label(ancova.results , "(a)")

@@

% Table S4
\clearpage
<<trait_pca, results="asis">>=
PCA.results <- round_df(PCA.results)
PCA.results <- embolden_top_values(PCA.results, row.select = c("dissection", "length", "pinna", "rhizome", "sla", "stipe", "width"))

print(xtable(PCA.results,
             label="supp_table:PCA",
             caption="PC loadings for principal components analyses (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia. For each PC, values of the three traits with the greatest loadings (absolute values) shown in bold."),
      sanitize.text.function = function (x) {x},
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      include.rownames=TRUE,
      caption.placement="top")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% set style
\bibliographystyle{newphyt}

% change the bibliography title
\renewcommand\refname{References}

% insert bibliography
\bibliography{references}

>>>>

\documentclass{article}

@@

@@

% Set unit and abbreviations formatting
\input{table_style}

@@

<<setup, echo=FALSE>>=
# setup knitr
library(knitr)
opts_knit$set(concordance=FALSE)
opts_chunk$set(echo=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(warning=FALSE)
@

@@

<<setup_tables>>=
library(xtable)
setwd(here::here())
source("bin/latex_functions.R")
# load results objects produced by running manuscript.Rnw
load("results.RData")
@

% Table 1
\clearpage

\begin{table}[ht]
\centering
\caption{Traits used in this study. Noncordate gametophyte morphotypes include ribbon, strap and filamentous.}
\begin{tabular}{ p{0.3\textwidth} p{0.2\textwidth} p{0.4\textwidth} }
\hline
 & Data type (unit) & Ecological Significance \\
\hline
Sporophyte Traits & & \\
\hspace{3ex}Stipe Length  & Continuous (cm)  & Shorter stipes compensate for low conductivity \citep{WatkinsJr.2010} \\
\hspace{3ex}Frond Length  & Continuous (cm)  & Smaller leaf size reduces evapotranspiration \citep{Vogel1968} \\
\hspace{3ex}Frond Width  & Continuous (cm)  & Smaller leaf size reduces evapotranspiration \citep{Vogel1968} \\
\hspace{3ex}Rhizome Diameter  & Continuous (cm)  & Scales with overall plant size (Creese et al. 2011) \\
\hspace{3ex}Frond Dissection  & Ordinal  & Simple laminae reduce evapotranspiration (Kluge and Kessler 2007) \\
\hspace{3ex}Pinna number & Integer & Scales with overall plant size (Creese et al. 2011) \\
\hspace{3ex}Specific Leaf Area & Continuous (m\textsuperscript{2} kg\textsuperscript{-1}) & Species in more stressful environments invest in thicker leaves \citep{Wright2004a} \\
Gametophyte Traits & & \\
\hspace{3ex}Morphotype & Binary (cordate vs. noncordate) & Noncordate morphology reduces drying rates by holding external water in crevices and folds \citep{WatkinsJr.2007,Pittermann2013} \\
\hspace{3ex}Gemmae & Binary (present/absent) & Enable asexual reproduction, thereby overcoming local extinction due to drying \citep{Farrar2008} \\
\hspace{3ex}Glands & Binary (present/absent) & Lipophilic exudates produced by glands may reduce transpiration rates \citep{Wollenweber2000} \\
\hspace{3ex}Hairs & Binary (present/absent) & Hairs reduce drying rates by holding external water and increasing boundary layer thickness \citep{WatkinsJr.2007} \\
\hline
\end{tabular}
\label{table:traits}
\end{table}

% Table 2
\clearpage

<<cont_phylosig, results="asis">>=
cont_phylosig.results <- round_df(cont_phylosig.results)
cont_phylosig.results <- embolden_p(cont_phylosig.results, col.select=c("k.pval","lambda.pval"))
part_label <- add_part_label(cont_phylosig.results, "(a)")

print(xtable(cont_phylosig.results,
             label="table:phylosig",
             caption="Phylogenetic signal in traits. (a) Continuous traits and (b) binary traits. Values of $\\lambda$ or $\\K >= 1$ indicate that traits are evolving according to Brownian Motion (BM); values of $\\lambda$ or \\K near 0 indicate random distribution of traits. For \\D, values of 0 (or less) indicate evolution according to (or more conserved than) BM; values near 1 indicate random trait distribution. \\pval\\textsubscript{Rnd}, result of testing that observed \\D is significantly different from 1; \\pval\\textsubscript{BM}, result of testing that observed \\D is significantly different from 0."),
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      sanitize.text.function = function (x) {x},
      include.rownames=TRUE,
      add.to.row = part_label,
      caption.placement="top")
@

<<binary_phylosig, results="asis">>=
binary_phylosig.results <- round_df(binary_phylosig.results)
binary_phylosig.results <- embolden_p(binary_phylosig.results, col.select=c("prob_random","prob_browian"))
part_label <- add_part_label(binary_phylosig.results, "(b)")

print(xtable(binary_phylosig.results,
             caption=NULL),
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      sanitize.text.function = function (x) {x},
      include.rownames=TRUE,      
      add.to.row = part_label)
@

% Table 3
\clearpage

<<sporo_PIC, results="asis">>=
PIC.results <- round_df(PIC.results)
PIC.results <- embolden_p(PIC.results, col.select=c("pval"))

print(xtable(PIC.results,
             label="table:PIC_sporo",
             caption="Results of phylogenetically independent contrasts analysis of quantitative (sporophyte) traits related to epiphytic growth. All trait contrasts were made by subtracting epiphytic values from terrestrial values; thus, a positive contrast indicates greater values for terrestrial clades. Significant differences between epiphytes and terrestrial clades in bold."),
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      sanitize.text.function = function (x) {x},
      include.rownames=TRUE,
      caption.placement="top")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%% Appendix Tables %%%%%%%%%%%%%%%%%%%%%%
% Add S to table number and reset counter
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

% Table S1
\clearpage

<<PGLMM, results="asis">>=
# read in results from running PGLMM on cluster, clean up dataframe
pglmm.results <- read.csv("data/mcmcglmm_habit_all_results.csv")
pglmm.results <- pglmm.results[pglmm.results$Effect == "habit", ]
pglmm.results$Effect <- NULL
pglmm.results$Model <- gsub(" ~ habit", "", pglmm.results$Model)
pglmm.results$Model <- tolower(pglmm.results$Model)
rownames(pglmm.results) <- pglmm.results$Model
pglmm.results$Model <- NULL

@@

% Table S2
\clearpage

<<binaryPGLMM, results="asis">>=
bpglmm.results <- round_df(bpglmm.results)
bpglmm.results <- embolden_p(bpglmm.results, col.select=c("sigmap","pvalue"))

@@

% Table S3
\clearpage

<<climate_ancova, results="asis">>=
ancova.results <- round_df(ancova.results)
ancova.results <- embolden_p(ancova.results, col.select=c("pval.intercept","pval.slope"))
part_label <- add_part_label(ancova.results , "(a)")

@@

% Table S4
\clearpage

<<trait_pca, results="asis">>=
PCA.results <- round_df(PCA.results)
PCA.results <- embolden_top_values(PCA.results, row.select = c("dissection", "length", "pinna", "rhizome", "sla", "stipe", "width"))

print(xtable(PCA.results,
             label="supp_table:PCA",
             caption="PC loadings for principal components analyses (PCA) of traits related to epiphytic growth in ferns from Moorea, French Polynesia. For each PC, values of the three traits with the greatest loadings (absolute values) shown in bold."),
      sanitize.text.function = function (x) {x},
      sanitize.colnames.function = custom_cols,
      sanitize.rownames.function = custom_rows,
      include.rownames=TRUE,
      caption.placement="top")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%% Refs in Tables %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{newphyt} 
\bibliography{references}
